Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}

### 2017-08-03 revision/review_response.Rnw

The authors executed the work on parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern Gulf of Mexico. The authors did not provide proper conceptual diagram of the model.

{\it The previous conceptual diagram was for CGEM that included both the biogeochemical components of FishTank and hydrodynamic components.  A new figure was created that inludes only the FishTank components.  }

The model is not presented in detail, it is necessary to give more details on original model.

{\it We have added Table 1 that inlcudes a full list and description of parameters that were evaluated for the sensitivity analysis.}

@@

Line 9 page 10. Selection or Selected?

{\it Changed to: `Heurestics for parameter selection also recognized...'}

Line 4 page 12. what are the default parameter conditions?. You should include a table with the default parameter values and parameter values under 50\% modified conditions.

{\it A new table (Table 1) was added for the full parameter set including names, long description, units, values, and the 50\% increase in values for the sensitivity analyses.}

### 2017-08-03 ident_manu.Rnw
parameter table added

The sensitivity of state variables to perturbations of all relevant parameters for the 108 equations was estimated using a five minute timestep with daily output from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006. Irrelevant parameters were removed for several reasons; parameters were not relevant for the \ac{zerod} model (i.e., hydrodynamic parameters), were considered physical constants, or had no effect given initial conditions.  Additionally, FishTank includes six phytoplankton and two zooplankton groups to add complexity in community structure and foodweb dynamics. To remove obvious redundancies, the sensitivity analyses were conducted using only one phytoplankton and one zooplankton group.  The final set that was evaluated included \Sexpr{length(p1z1pars)} parameters that were further grouped into one of six categories based on applicable biogeochemical components of the model: optics ($n = $ \Sexpr{nparms$Optics} parameters), organic matter (\Sexpr{nparms$`Organic Matter`}), phytoplankton (\Sexpr{nparms$Phytoplankton}), temperature (\Sexpr{nparms$Temperature}), and zooplankton (\Sexpr{nparms$Zooplankton}).  A full description of the model parameters is available as an appendix in \citet{Lehrter17}.

@@

A local sensitivity analysis was performed by evaluating the change in state variables following perturbation of each parameter from its original value \citep{Soetaert10,Camacho14b,RDCT17}. This approach is described as a First Order Variance Analysis that linearly propagates changes from model parameters to model predictions \citep{Camacho14b}. Parameters were individually perturbed by a 50\% increase of the original values and sensitivity $S$ was estimated for each time step $i$ given a change for parameter $j$ as:

@@

@@

% model schematic
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{figs/fishmod.pdf}
\caption{Conceptual representation of biogeochemical components included in FishTank (complete equations in \citealt{Eldridge10}). The \acl{cgem} couples FishTank with a hydrodynamic model that includes advection, mixing, and dispersion.}
\label{fig:mod_schem}
\end{figure}

% combination example
<<combnex, eval = T, echo = F, fig.cap = 'Examples of unique parameter combinations from different parameter sets and number of selected parameters.  The number of combinations are shown for increasing numbers of selected parameters from the total in the set, where 50 parameter sets are shown each with one through 50 total parameters. Note that the number of unique combinations is shown as the natural-log.', fig.height = 5, fig.width = 5, out.width = '0.6\\textwidth'>>=

@@

% sensitivity heat map
<<sensalltile, fig.cap = 'Sensitivity values (L1, \\cref{l1}) of all state variables to changes in a 50\\% increase in parameter values. Parameters are grouped by category: optics, organic matter, phytoplankton, zooplankton, temperature, and zoplankton.  See \\cref{tab:dosens} for L1 values for \\ac{do} and \\cref{tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens} for the other state variables.', fig.height = 6, fig.width = 4, out.width = '0.7\\textwidth', eval = T, echo = F, warning = F, message = F>>=
# initial conditions
data(ini_eval)
data(struct_eval)

>>>>

The sensitivity of state variables to perturbations of all relevant parameters for the 108 equations was estimated using a five minute timestep with daily output from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006. Irrelevant parameters were removed for several reasons; parameters were not relevant for the \ac{zerod} model (i.e., hydrodynamic parameters), were considered physical constants, or had no effect given initial conditions.  Additionally, FishTank includes six phytoplankton and two zooplankton groups to add complexity in community structure and foodweb dynamics. To remove obvious redundancies, the sensitivity analyses were conducted using only one phytoplankton and one zooplankton group.  The final set that was evaluated included \Sexpr{length(p1z1pars)} parameters that were further grouped into one of six categories based on applicable biogeochemical components of the model: optics ($n = $ \Sexpr{nparms$Optics} parameters), organic matter (\Sexpr{nparms$`Organic Matter`}), phytoplankton (\Sexpr{nparms$Phytoplankton}), temperature (\Sexpr{nparms$Temperature}), and zooplankton (\Sexpr{nparms$Zooplankton}) (\cref{tab:parmtab}).  A full description of the model parameters is available as an appendix in \citet{Lehrter17}.

@@

A local sensitivity analysis was performed by evaluating the change in state variables following perturbation of each parameter from its original value \citep{Soetaert10,Camacho14b,RDCT17}. This approach is described as a First Order Variance Analysis that linearly propagates changes from model parameters to model predictions \citep{Camacho14b}. Parameters were individually perturbed by a 50\% increase of the original values (\cref{tab:parmtab}) and sensitivity $S$ was estimated for each time step $i$ given a change for parameter $j$ as:

@@

% parameter info
<<parmtab, eval = T, cache = F>>=
parmtab(tabsize = 'tiny')
@

@@

% model schematic
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{figs/fishmod.pdf}
\caption{Conceptual representation of biogeochemical components included in FishTank (complete equations in \citealt{Eldridge10}). The \acl{cgem} couples FishTank with a hydrodynamic model that includes advection, mixing, and dispersion.}
\label{fig:fishmod}
\end{figure}

% combination example
<<combnex, eval = T, echo = F, fig.cap = 'Examples of unique parameter combinations from different parameter sets and number of selected parameters.  The number of combinations are shown for increasing numbers of selected parameters from the total in the set, where 50 parameter sets are shown each with one through 50 total parameters. Note that the number of unique combinations is shown as the natural-log.', cache = TRUE, fig.height = 5, fig.width = 5, out.width = '0.6\\textwidth'>>=

@@

% sensitivity heat map
<<sensalltile, fig.cap = 'Sensitivity values (L1, \\cref{l1}) of all state variables to changes in a 50\\% increase in parameter values. Parameters are grouped by category: optics, organic matter, phytoplankton, zooplankton, temperature, and zoplankton.  See \\cref{tab:dosens} for L1 values for \\ac{do} and \\cref{tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens} for the other state variables.', fig.height = 6, fig.width = 4, out.width = '0.7\\textwidth', eval = T, cache = TRUE, echo = F, warning = F, message = F>>=
# initial conditions
data(ini_eval)
data(struct_eval)

### 2017-08-03 ident_manu.tex
parameter table added

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Wed Aug 2 13:16:06 2017 -0500, \href{https://github.com/fawda123/identifiability/commit/5cd8b68899d08c2c8f96177fe72e018bd764f96a}{5cd8b68899d08c2c8f96177fe72e018bd764f96a}\newline Acronyms: \acfi{chla}; \acfi{cgem}; \acfi{do}; \acfi{dom}; \acfi{gom}; \acfi{lcs}; \acfi{marb}; \acfi{pom}; \acfi{rmse}; \acfi{zerod}}

@@

The sensitivity of state variables to perturbations of all relevant parameters for the 108 equations was estimated using a five minute timestep with daily output from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006. Irrelevant parameters were removed for several reasons; parameters were not relevant for the \ac{zerod} model (i.e., hydrodynamic parameters), were considered physical constants, or had no effect given initial conditions.  Additionally, FishTank includes six phytoplankton and two zooplankton groups to add complexity in community structure and foodweb dynamics. To remove obvious redundancies, the sensitivity analyses were conducted using only one phytoplankton and one zooplankton group.  The final set that was evaluated included 51 parameters that were further grouped into one of six categories based on applicable biogeochemical components of the model: optics ($n = $ 4 parameters), organic matter (12), phytoplankton (22), temperature (2), and zooplankton (11).  A full description of the model parameters is available as an appendix in \citet{Lehrter17}.

@@

@@

% model schematic
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{figs/fishmod.pdf}
\caption{Conceptual representation of biogeochemical components included in FishTank (complete equations in \citealt{Eldridge10}). The \acl{cgem} couples FishTank with a hydrodynamic model that includes advection, mixing, and dispersion.}
\label{fig:mod_schem}
\end{figure}

>>>>

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Wed Aug 2 17:20:03 2017 -0500, \href{https://github.com/fawda123/identifiability/commit/a49e333c59176457f6c598684d90ef37895bdafb}{a49e333c59176457f6c598684d90ef37895bdafb}\newline Acronyms: \acfi{chla}; \acfi{cgem}; \acfi{do}; \acfi{dom}; \acfi{gom}; \acfi{lcs}; \acfi{marb}; \acfi{pom}; \acfi{rmse}; \acfi{zerod}}

@@

The sensitivity of state variables to perturbations of all relevant parameters for the 108 equations was estimated using a five minute timestep with daily output from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006. Irrelevant parameters were removed for several reasons; parameters were not relevant for the \ac{zerod} model (i.e., hydrodynamic parameters), were considered physical constants, or had no effect given initial conditions.  Additionally, FishTank includes six phytoplankton and two zooplankton groups to add complexity in community structure and foodweb dynamics. To remove obvious redundancies, the sensitivity analyses were conducted using only one phytoplankton and one zooplankton group.  The final set that was evaluated included 51 parameters that were further grouped into one of six categories based on applicable biogeochemical components of the model: optics ($n = $ 4 parameters), organic matter (12), phytoplankton (22), temperature (2), and zooplankton (11) (\cref{tab:parmtab}).  A full description of the model parameters is available as an appendix in \citet{Lehrter17}.

@@

% parameter info
{\tiny
\caption{FishTank parameters evaluated for sensitivity and identifiability. Sensitivitity analyses were based on a 50\% increase (Value $\cdot$ 1.5) from the initial value of each parameter. Dashed units are dimensionless. Parameters are grouped by categories as optics, temperature, phytoplankton, zooplankton, and organic matter.  A full description of the model structure, equations, and parameters is described in \citet{Lehrter17}.\label{tab:parmtab}} 
\multicolumn{1}{l}{Description}&\multicolumn{1}{c}{Parameter}&\multicolumn{1}{c}{Units}&\multicolumn{1}{c}{Value}&\multicolumn{1}{c}{Value $\cdot$ 1.5}\tabularnewline
{\bfseries Optics}&&&&\tabularnewline
~~Chla specific absorption at 490 nm&\textit{astar490}&m$^{-1}$ (mg Chla m$^{-3}$)$^{-1}$&$0.04$&$0.06$\tabularnewline
~~OMA specific absorption at 490 nm&\textit{astarOMA}&m$^{-1}$ (mg OMA m$^{-3}$)$^{-1}$&$0.1$&$0.15$\tabularnewline
~~OMZ specific absorption at 490 nm&\textit{astarOMZ}&m$^{-1}$ (mg OMZ m$^{-3}$)$^{-1}$&$0.1$&$0.15$\tabularnewline
~~sinking rate&\textit{sink CDOM}&m d$^{-1}$&$0$&$0$\tabularnewline
{\bfseries Temperature}&&&&\tabularnewline
~~Optimum temperature for growth&\textit{Tref(nospA+nospZ)$_{p1}$}&C&$22$&$33$\tabularnewline
~~Optimum temperature for growth&\textit{Tref(nospA+nospZ)$_{p1}$}&C&$22$&$33$\tabularnewline
{\bfseries Phytoplankton}&&&&\tabularnewline
~~initial slope of photosynthesis v irradiance&\textit{alpha}&10$^{-16}$ cm$^2$ s quanta$^{-1}$ d$^{-1}$&$8.42\times 10^{-17}$&$1.26\times 10^{-16}$\tabularnewline
~~coefficient for non-limiting nutrient&\textit{aN}&-&$1$&$1.5$\tabularnewline
~~phytoplankton threshold for grazing&\textit{Athresh}&10$^7$ cells m$^{-3}$&$1.72\times 10^{8}$&$2.58\times 10^{8}$\tabularnewline
~~edibility vector for Z1&\textit{ediblevector(Z1)}&-&$0.25$&$0.38$\tabularnewline
~~edibility vector for Z2&\textit{ediblevector(Z2)}&-&$0.25$&$0.38$\tabularnewline
~~half-saturation constant for N&\textit{Kn}&mmol m$^{-3}$&$4.51$&$6.76$\tabularnewline
~~half-saturation constant for P&\textit{Kp}&mmol m$^{-3}$&$2.86$&$4.29$\tabularnewline
~~Qn constant for Flynn growth model&\textit{KQn}&-&$5$&$7.5$\tabularnewline
~~Qp constant for Flynn growth model&\textit{KQp}&-&$0.2$&$0.3$\tabularnewline
~~half-saturation constant for Si uptake&\textit{Ksi}&mmol m$^{-3}$&$4.51$&$6.76$\tabularnewline
~~mortality coefficient&\textit{mA}&d$^{-1}$&$0.1$&$0.15$\tabularnewline
~~phytoplankton carbon/cell&\textit{Qc}&10$^{-7}$ mmol C cell$^{-1}$&$1.35\times 10^{-6}$&$2.03\times 10^{-6}$\tabularnewline
~~minimum N cell-quota&\textit{QminN}&10$^{-9}$ mmol N cell$^{-1}$&$6.08\times 10^{-9}$&$9.12\times 10^{-9}$\tabularnewline
~~minimum P cell-quota&\textit{QminP}&10$^{-9}$ mmol P cell$^{-1}$&$6.19\times 10^{-10}$&$9.29\times 10^{-10}$\tabularnewline
~~phytoplankton basal respiration coefficient&\textit{respb}&d$^{-1}$&$0.02$&$0.03$\tabularnewline
~~phytoplankton growth respiration coefficient&\textit{respg}&-&$0.1$&$0.15$\tabularnewline
~~sinking rate of phytoplankton cells&\textit{sink A}&m d$^{-1}$&$1.49$&$2.23$\tabularnewline
~~maximum growth rate&\textit{umax}&d$^{-1}$&$0.41$&$0.62$\tabularnewline
~~N-uptake rate measured at umax&\textit{vmaxN}&10$^{-8}$ mmol cell$^{-1}$ d$^{-1}$&$4.1\times 10^{-8}$&$6.15\times 10^{-8}$\tabularnewline
~~P-uptake rate measured at umax&\textit{vmaxP}&10$^{-8}$ mmol cell$^{-1}$ d$^{-1}$&$2.68\times 10^{-8}$&$4.02\times 10^{-8}$\tabularnewline
~~Si-uptake rate measured at umax&\textit{vmaxSi}&10$^{-8}$ mmol cell$^{-1}$ d$^{-1}$&$4.1\times 10^{-8}$&$6.15\times 10^{-8}$\tabularnewline
~~phytoplankton volume/cell&\textit{volcell}&$\mu$m$^3$&$3.37\times 10^{4}$&$5.05\times 10^{4}$\tabularnewline
{\bfseries Zooplankton}&&&&\tabularnewline
~~assimilation efficiency as a fraction of ingestion&\textit{Zeffic}&-&$0.4$&$0.6$\tabularnewline
~~half saturation coefficient for grazing&\textit{ZKa}&$\mu$m$^3$ m$^{-3}$&$1.12\times 10^{12}$&$1.68\times 10^{12}$\tabularnewline
~~quadratic mortality constant&\textit{Zm}&m$^6$ ind$^{-2}$ d$^{-1}$&$7.2\times 10^{-4}$&$0$\tabularnewline
~~zooplankton carbon/individual&\textit{ZQc}&mmol C ind$^{-1}$&$3.13\times 10^{-4}$&$4.7\times 10^{-4}$\tabularnewline
~~zooplankton nitrogen/individual&\textit{ZQn}&mmol N ind$^{-1}$&$6.95\times 10^{-5}$&$1.04\times 10^{-4}$\tabularnewline
~~zooplankton phosphorus/individual&\textit{ZQp}&mmol P ind$^{-1}$&$3.77\times 10^{-6}$&$5.66\times 10^{-6}$\tabularnewline
~~zooplankton biomass-dependent respiration factor&\textit{Zrespb}&d$^{-1}$&$0.1$&$0.15$\tabularnewline
~~zooplankton growth-dependent respiration factor&\textit{Zrespg}&-&$0.2$&$0.3$\tabularnewline
~~proportion of phytoplankton lost to sloppy feeding&\textit{Zslop}&-&$0.25$&$0.38$\tabularnewline
~~maximum growth rate of zooplankton&\textit{Zumax}&$\mu$m$^3$ ind$^{-1}$ d$^{-1}$&$9.45\times 10^{7}$&$1.42\times 10^{8}$\tabularnewline
~~zooplankton volume/individual&\textit{Zvolcell}&$\mu$m$^3$ ind$^{-1}$&$2.98\times 10^{7}$&$4.47\times 10^{7}$\tabularnewline
{\bfseries Organic Matter}&&&&\tabularnewline
~~turnover rate for OM1A and OM1Z&\textit{KG1}&y$^{-1}$&$50$&$75$\tabularnewline
~~turnover rate for OM2A and OM2Z&\textit{KG2}&y$^{-1}$&$50$&$75$\tabularnewline
~~decay rate of CDOM&\textit{KGcdom}&d$^{-1}$&$0.01$&$0.02$\tabularnewline
~~NH4 rate constant for nitrification&\textit{KNH4}&mmol m$^{-3}$&$1$&$1.5$\tabularnewline
~~half-saturation concentration for denitrification&\textit{KNO3}&mmol m$^{-3}$&$10$&$15$\tabularnewline
~~half-saturation concentration for O2 utilization&\textit{KO2}&mmol m$^{-3}$&$10$&$15$\tabularnewline
~~O2 concentration that inhibits denitrification&\textit{KstarO2}&mmol m$^{-3}$&$10$&$15$\tabularnewline
~~maximum rate of nitrification per day&\textit{nitmax}&mmol m$^{-3}$ d$^{-1}$&$0.52$&$0.78$\tabularnewline
~~phytoplankton sinking rate&\textit{sink OM1\_A}&m d$^{-1}$&$10$&$15$\tabularnewline
~~zooplankton sinking rate&\textit{sink OM1\_Z}&m d$^{-1}$&$10$&$15$\tabularnewline
~~phytoplankton sinking rate&\textit{sink OM2\_A}&m d$^{-1}$&$0$&$0$\tabularnewline
~~zooplankton sinking rate&\textit{sink OM2\_Z}&m d$^{-1}$&$0$&$0$\tabularnewline

@@

% model schematic
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{figs/fishmod.pdf}
\caption{Conceptual representation of biogeochemical components included in FishTank (complete equations in \citealt{Eldridge10}). The \acl{cgem} couples FishTank with a hydrodynamic model that includes advection, mixing, and dispersion.}
\label{fig:fishmod}
\end{figure}
